Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy

### 2021-08-04 manuscript.tex
just working on 3.7 and 3.8
JACS resubmission

%\newcommand*{\rood}[1]{{\color{red}{#1}}}
%\newcommand*{\blauw}[1]{\textcolor{blue}{#1}}
\newcommand*{\blauw}[1]{{#1}}

%\newcommand*{\addref}[1]{\textcolor{red}{\{ADD REF: #1\}}}
%\newcommand*{\noter}[1]{\textcolor{red}{[[#1]]}}		% notes 
%\newcommand*{\noter}[1]{}					% notes off
%\newcommand*{\noteb}[1]{\textcolor{blue}{[[#1]]}}		% notes

@@

Over the last couple of decades, DNA has proven to be much more than a vessel for genetic information. From sensing to computing to directed self-assembly, the programmable and predictable nature of DNA has unlocked numerous unforeseen nanotechnology applications.\citep{Seeman2017DNANanotechnology, Adleman1994MolecularProblems, Rothemund2006FoldingPatterns, Gu2010ALine} Recently, single molecule localization techniques have exploited the rapid and transient binding of short DNA oligomers in order to achieve super-resolution microscopy and optical multiplexing. \cite{Schnitzbauer2017Super-resolutionDNA-PAINT, Shah2019, Strauss2020UpDNA-PAINT} Predictive understanding of the sequence-dependent thermodynamics of DNA hybridization/dehybridization -- the assembly/disassembly of a DNA duplex from two single strands -- has underpinned the rational design of DNA oligomer sequences for nanotechnology applications, where sequence-dependent nearest-neighbor models can accurately account for mismatched pairs, dangling ends, and other non-native bonding effects.\citep{SantaLucia1998AThermodynamics, Santalucia2004TheMotifs} Secondary DNA structures such as hairpins and G-quadruplexes have also been studied in depth and leveraged for nanotechnology applications.\citep{Tsukanov2013DetailedOrigami, Mosayebi2014TheFormation, Mergny2019DNANanotechnology} Predictive models of the dynamical, as opposed to purely thermodynamical, behaviors of DNA have become increasingly important in developing technologies such as DNA-PAINT (DNA Points Accumulation For Imaging In Nanoscale Topography), but these technologies have outpaced our fundamental understanding of the dynamics themselves. \citep{Deluca2020DynamicDevices, Cordes2010SensingTransfer, Naimark2020DNADevelopment} Many experimental and computational studies have investigated DNA dynamical phenomena over picosecond to millisecond time scales.\citep{Yin2011KineticsHybridization, Xiao2019, Hinckley2014Coarse-grainedEffects, Sanstead2016, Porschke1971CooperativeTransition} Kinetic models have been developed for particular DNA processes such as toehold exchanges and optical barcoding \citep{Zhang2009ControlExchange, Shah2019ProgrammingFingerprinting} and supervised machine learning techniques have been combined with experimental measurements to predict the on/off rates as a function of sequence. \citep{Schickinger2018TetheredHelices, Zhang2018PredictingSequence, Shah2019}. A comprehensive understanding of the full dynamical landscape of hybridization/dehybridization accounting for the sequence-dependent metastable states and association/dissociation pathways remains lacking and fundamental questions remain unresolved. For example, it remains unclear the extent to which hybridization of short DNA oligomers largely proceeds in a conventionally assumed ``all-or-nothing'' fashion or if long-lived metastable states facilitate the transition.\citep{Araque2016LatticeCooperativity, Sikora2013ModelingIntermediates, Wyer2014KineticsAT-tracts, Sanstead2016, Sanstead2018DirectDehybridization, Xiao2019} Out-of-register ``shifted'' base paired structures \citep{Hinckley2014Coarse-grainedEffects, Maciejczyk2014DNAModel, Araque2016LatticeCooperativity, Xiao2019} and frayed structures \citep{Zgarbova2014BaseRNA, Nonin1995TerminalFraying, Nikolova2012ProbingSimulations, Andreatta2006UltrafastHelix} stand as candidates for metastable states with the potential to mediate substantial deviations from all-or-nothing behavior, but the degree to which these states are kinetically relevant is difficult to determine experimentally and is likely to be highly sequence-dependent. The development of predictive models and design rules with which to engineer DNA strands with tailored hybridization/dehybridization kinetics and pathways is vital to advancing rational design of DNA strands for nanotechnology applications and is also of importance in understanding fundamental biological processes such as transcription and gene regulation.

Our understanding of hybridization dynamics is built upon decades of experiments -- such as temperature-jump, salt-jump, pH-jump, and other perturbative methods -- that rapidly stimulate DNA and monitor relaxation to a new equilibrium.\citep{Morrison1993SensitiveSolution, Wetmur1968KineticsDNA, Craig1971RelaxationOligonucleotides, Porschke1973ThermodynamicsPairs, Williams1989LaserDGCATGC, Narayanan2012ExploringMixing, Chen2007InfluenceHybridization, Sanstead2018DirectDehybridization} More recently, single molecule diffusion and tethered multifluorophore assays have facilitated analyses under equilibrium conditions, but these results can be hampered by slow data collection rates and fluorescent tags effects on strand dynamics, particularly for shorter oligomers.\citep{Liu20173DSolution,  Schickinger2018TetheredHelices, Chen2008Base-by-baseSpectroscopy, Dupuis2013Single-moleculeHelices} A number of computational modeling approaches have also been employed to provide molecular-level resolution of hybridization. Simplified lattice models can recapitulate the essential aspects of the hybridization pathways but lack the realism of continuous space representations \citep{Araque2016LatticeCooperativity, Phys2019}. The long time scales associated with hybridization/dehybridization events place them outside the reach of unbiased all-atom molecular dynamics simulations,\citep{Hinckley2014Coarse-grainedEffects} but they can be observed by employing enhanced sampling techniques\citep{Piana2007AtomisticTransition, Zerze2021ThermodynamicsSimulations, Hinckley2013AnHybridization, Schmitt2013ExploringSurface, Sambriski2009, Sambriski2009SequencePathways, Hoefert2011MolecularOligonucleotides,Romano2013DNADependence, Hinckley2014Coarse-grainedEffects} or by using elevated temperature or denaturing solvent concentrations to induce one-way dissociation events.\citep{Wong2008TheSimulations, Perez2010Real-timeUnfolding} The effect of the applied bias upon the thermodynamics can be rigorously corrected for using standard reweighting techniques, but approaches to rigorously correct the kinetics, particularly under the conditions of high bias necessary for good sampling, are in their infancy.~\citep{Prinz2011OptimalDynamics, Chodera2011DynamicalTemperatures, Stelzl2017DynamicSimulations, Donati2017GirsanovModels, Donati2018GirsanovSimulations, Quer2018ANCOORDINATES} A number of coarse-grained DNA force fields have been developed that enable direct observation of these events over microsecond time scales via unbiased coarse-grained molecular dynamics simulations, \citep{Romano2013DNADependence, Hinckley2013AnHybridization, Maciejczyk2014DNAModel, Markegard2015, Dans2016MultiscaleDNA} which, up to an acceleration factor associated with the smoothing of the underlying free energy landscape inherent to the coarse-graining procedure, can preserve a faithful model of the unbiased dynamics and associated pathways. These models have previously been used to study biological phenomena such as nucleosome dynamics\citep{Lequieu2016Tension-dependentUnwrapping, Lequieu2017InSliding} and transcription factor binding \cite{Terakawa2015P53Simulations, Tan2018DynamicData} as well as nanotechnology applications such as strand displacement \citep{Srinivas2013OnDisplacement, Haley2020DesignDisplacement} and DNA origami \citep{Snodin2019Coarse-grainedOrigami, Doye2020TheOrigami}.

@@

All calculations were performed using the LAMMPS simulation package (\url{http://lammps.sandia.gov}) in accordance with best practices for the 3SPN.2 model. \citep{Plimpton1997FastDynamics} A single pair of self-complementary sequences were placed in a cubic periodic box with side length 7.8 nm corresponding to a single-strand concentration of 7 mol/L. Solvent effects were modeled implicitly by employing Langevin dynamics \citep{DunwegB.Paul1991BrownianNumbers., Bussi2007AccurateDynamics} with an experimentally motivated per-site friction coefficient of 9.94 $\times$ 10$^{-11}$ m$^2$/s. \citep{Nkodo2001DiffusionElectrophoresis, Hinckley2013AnHybridization} We specified a 240 mM implicit salt concentration and treated electrostatic interactions using the Debye-H{\"u}ckel with a 5 nm cutoff radius. \citep{Debye1923ZurElektrolyte} Simulations were performed in the NVT ensemble employing a Langevin thermostat. \citep{Schneider1978Molecular-dynamicsTransitions}. Each sequence was simulated at its melting temperature -- AT-all: 309 K, GC-end: 317 K, GC-core: 324 K, GC-mix: 324 K -- in order to maximize the number of spontaneous transitions between dissociated and hybridized states. Melting temperatures for each sequence were determined empirically by comparing the ratio of hybridized to dissociated populations over a 10 K temperature ramp centered on the nearest neighbor (NN) model predicted melting temperature,~\citep{SantaLucia1998AThermodynamics, Santalucia2004TheMotifs} and selecting the temperature at which the hybridized and dissociated populations were approximately equal. The Langevin equations of motion were integrated using the scheme of Bussi and Parrinello \citep{Bussi2007AccurateDynamics} with a 20 fs integration time step. We performed 40 independent simulations for each of the four sequences with half of the runs initialized from the hybridized state and half from the dissociated state. The initial hybridized state was defined based on the crystal structure coordinates of Arnott et al.~\citep{Arnott1976CRCBiology} The dissociated state was generated from the hybridized state by displacing one strand away from the other by 1 nm in each of the $x$, $y$, and $z$ directions. The magnitude of this displacement did not affect the results so long as all native Watson-Crick (WC) bonds were completely broken. Initial bead velocities were assigned from a Maxwell-Boltzmann distribution at the temperature of interest. Each simulation was conducted for 26 $\mu$s and frames saved to disc every 100 ps. Each simulation required $\sim$24 CPU-hours on 28$\times$Intel E5-2680v4 CPU cores. The first 1 $\mu$s of each run was discarded for equilibration providing us with 40$\times$25 $\mu$s = 1 ms of simulation data for each sequence, during which time we observed 55-100 hybridization/dehybridization events.

@@

Markov state models (MSMs) are a powerful approach to infer long-time kinetic models from short molecular simulation trajectories \citep{Sengupta2019AutomatedSelf-assembly, Pande2010EverythingAsk, Chodera2014MarkovDynamics, Husic2018MarkovScience, Wehmeyer2019IntroductionSoftware} that we employ in this work to construct high-resolution sequence-dependent kinetic models of DNA hybridization and dissociation. In brief, a MSM extracts from simulation trajectories an ensemble of long-lived metastable macrostates, their equilibrium occupancy populations, and the equilibrium probability fluxes between them. In this manner, they provide an interpretable and predictive model of the system thermodynamics and kinetics. MSMs have recently been implemented to study the hybridization mechanism of one particular 14-mer DNA oligomer, but determining the sequence-dependent kinetics and mechanisms was not the focus of this study.\citep{Jin2019} An energy disconnectivity graph-based approach was used to interrogate the differences in hybridization rates and mechanisms between GGGGGG and GCGCGC hexamers to reveal strong deviations from ``all-or-nothing'' behaviors and the importance of zippering and slithering mechanisms.~\citep{Xiao2019} Kinetic models were constructed not from the dynamical trajectories of the molecular model, but by estimating rate constants between local minima using a transition state theory approximation. A recent application of MSMs to the long-time dynamics of short RNA oligonucleotides revealed stacking time scales to be highly sequence dependent.~\citep{Pinamonti2017} In this work, MSMs were constructed for each of the four DNA sequences from the 40$\times$25 $\mu$s simulation trajectories following a four-step protocol detailed in Ref.~\citep{Sidky2019High-ResolutionVAMPnets}: (i) trajectory featurization, (ii) dimensionality reduction, (iii) microstate clustering and microstate transition matrix inference, (iv) macrostate clustering and macrostate transition matrix inference. Calculations were performed using the PyEMMA software package. \citep{Scherer2015PyEMMAModels}

@@

We first sought to demonstrate that the coarse-grained 3SPN.2 model accurately recapitulates the experimentally observed kinetics of DNA oligomer hybridization/dissociation by validating our computational predictions against temperature-jump infrared (T-jump IR) experiments. We conducted T-jump IR experiments for each DNA sequence as a function of temperature and extracted the ``slow'' $k_d^\mathrm{slow}$ (10-30 $\mu$s) and ``fast'' $k_d^\mathrm{fast}$ (70-100 ns) rates estimated from the T-jump response. The slow response has previously been attributed to duplex dissociation on microsecond time scales induced by the rapid heating of the initially hybridized duplex.~\citep{Sanstead2016, Sanstead2018DirectDehybridization} The fast response has been assigned to terminal base pair fraying,~\citep{Sanstead2016, Sanstead2018DirectDehybridization} with this process corresponding to a relatively complex dynamical process that can span time scales from picoseconds to microseconds.~\cite{Nonin1995TerminalFraying, Nikolova2012ProbingSimulations, Andreatta2006UltrafastHelix, Galindo-Murillo2015ConvergenceDGCACGAACGAACGAACGC} All-atom simulations suggest that frayed ends can assume misaligned WC bonds, base-sugar hydrogen bonds, and terminal stacked conformations.~\citep{Pinamonti2019TheModels, Zgarbova2014BaseRNA}

To computationally mimic the T-jump process in our 3SPN.2 simulations, we conducted 1 $\mu$s simulations of an initially hybridized DNA duplex over a range of temperatures and monitored its structural relaxation. We performed 120 independent simulations for each DNA sequence at each temperature, and from these extracted computational estimates of $k_d^\mathrm{slow}$ and $k_d^\mathrm{fast}$ (\blauw{Fig.~S4}). First, we tracked the slow response corresponding to duplex dissociation in our simulations by compiling the distribution of times at which both of the central base pairs first separate to a distance of 1.3 nm starting from an initial fully hybridized duplex. This cutoff was selected as the distance beyond with the strands are effectively non-interacting and defines the dissociated state. We extracted our computational estimate of $k_d^\mathrm{slow}$ by fitting a decaying exponential to the fraction of hybridized sequences as a function of time $f_\mathrm{hybridized}(t) = \exp(-k_d^\mathrm{slow}t)$. Second, we tracked the fast response corresponding to terminal base pair fraying by compiling the distribution of times at which either of the terminal base pairs first separated to a distance of 1.3 nm, corresponding to a complete breakage of the WC interaction. We extracted our computational estimate of $k_d^\mathrm{fast}$ through a decaying exponential fit to the fraction of unfrayed sequences as a function of time $f_\mathrm{unfrayed}(t) = \exp(-k_d^\mathrm{fast}t)$.

We present in Fig.~\ref{fig:relaxation-comparison} a comparison of $k_d^\mathrm{slow}$ and $k_d^\mathrm{fast}$ estimated by computation and experiment. Although the 3SPN.2 model was not directly fitted against kinetic data,~\cite{Hinckley2013AnHybridization} its predictions of sequence dependent T-jump relaxation rates are, within a systematic scaling factor in time and systematic shift in temperature, in good agreement with observed experimental trends. It is well-known that the smoothing of the underlying free energy landscape induced by coarse-graining artificially accelerates the kinetics of coarse-grained molecular simulations and that different degrees of freedom may be accelerated by different factors.~\citep{Marrink2007TheSimulations, Fritz2011MultiscaleDynamics, Marrink2013PerspectiveModel} We find that the simulated slow responses corresponding to center-of-mass translation of the strands during dissociation of the duplex is $\sim$10$\times$ accelerated relative to experiment, whereas the fast responses corresponding to fraying of the terminal bases is $\sim$120$\times$ accelerated. We apply these sequence-independent scaling factors to our reported computational values in Fig.~\ref{fig:relaxation-comparison}. Although the 3SPN.2 model reproduces melting temperatures relatively well, we observed a systematic 4 K under-prediction relative to experiment and so we apply a universal (+4) K corrective calibration to our computational results. We note that these empirical calibration factors to the 3SPN.2 predictions are applied only for the purposes of making an experimental comparison, but acknowledge that there are uncertainties introduced by assuming that the equilibrium dynamics at fixed temperature can be compared directly to relaxation kinetics following a 15$^\circ$C T-jump. The computational time scales and melting temperatures reported in the remainder of the paper are not corrected since we are only interested in the relative trends in the behaviors of the four sequences.

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=\textwidth]{Fig1.pdf}
        \caption{Experimental measurements and computational predictions of slow and fast at T-jump IR responses. Results are reported in terms of the final T-jump temperature. (a) The experimental and simulated slow rate constants $k_d^\mathrm{slow}$ corresponding to duplex dissociation over long time scales. (b) The experimental and simulated fast rate constant $k_d^\mathrm{fast}$ corresponding to terminal base-pair fraying on short time scales. The simulation results are corrected by a sequence-independent scaling factor that corrects for a 10$\times$ acceleration of the slow dissociation dynamics and 120$\times$ acceleration of the fast fraying dynamics. The simulated temperature in all cases is subjected to a (+4) K corrective calibration to account for an observed systematic under-prediction of the melting temperature by the 3SPN.2 model}
        \label{fig:relaxation-comparison}
	\end{center}
\end{figure}

>>>>

\newcommand*{\rood}[1]{{\color{red}{#1}}}
%\newcommand*{\blauw}[1]{\textcolor{blue}{#1}}
\newcommand*{\blauw}[1]{{#1}}

\newcommand*{\addref}[1]{\textcolor{red}{\{ADD REF: #1\}}}
\newcommand*{\noter}[1]{\textcolor{red}{[[#1]]}}		% notes 
%\newcommand*{\noter}[1]{}					% notes off
\newcommand*{\noteb}[1]{\textcolor{blue}{[[#1]]}}		% notes

@@

Over the last couple of decades, DNA has proven to be much more than a vessel for genetic information. From sensing to computing to directed self-assembly, the programmable and predictable nature of DNA has unlocked numerous unforeseen nanotechnology applications.\citep{Seeman2017DNANanotechnology, Adleman1994MolecularProblems, Rothemund2006FoldingPatterns, Gu2010ALine} Recently, single molecule localization techniques have exploited the rapid and transient binding of short DNA oligomers in order to achieve super-resolution microscopy and optical multiplexing. \cite{Schnitzbauer2017Super-resolutionDNA-PAINT, Shah2019, Strauss2020UpDNA-PAINT} Predictive understanding of the sequence-dependent thermodynamics of DNA hybridization/dehybridization -- the assembly/disassembly of a DNA duplex from two single strands -- has underpinned the rational design of DNA oligomer sequences for nanotechnology applications, where sequence-dependent nearest-neighbor models can accurately account for mismatched pairs, dangling ends, and other non-native bonding effects.\citep{SantaLucia1998AThermodynamics, Santalucia2004TheMotifs} Secondary DNA structures such as hairpins and G-quadruplexes have also been studied in depth and leveraged for nanotechnology applications.\citep{Tsukanov2013DetailedOrigami, Mosayebi2014TheFormation, Mergny2019DNANanotechnology} Predictive models of the dynamical, as opposed to purely thermodynamical, behaviors of DNA have become increasingly important in developing technologies such as DNA-PAINT (DNA Points Accumulation \rood{f}or Imaging \rood{i}n Nanoscale Topography), but these technologies have outpaced our fundamental understanding of the dynamics themselves. \citep{Deluca2020DynamicDevices, Cordes2010SensingTransfer, Naimark2020DNADevelopment} Many experimental and computational studies have investigated DNA dynamical phenomena over picosecond to millisecond time scales.\citep{Yin2011KineticsHybridization, Xiao2019, Hinckley2014Coarse-grainedEffects, Sanstead2016, Porschke1971CooperativeTransition} Kinetic models have been developed for particular DNA processes such as toehold exchanges and optical barcoding \citep{Zhang2009ControlExchange, Shah2019ProgrammingFingerprinting} and supervised machine learning techniques have been combined with experimental measurements to predict the on/off rates as a function of sequence. \citep{Schickinger2018TetheredHelices, Zhang2018PredictingSequence, Shah2019}. A comprehensive understanding of the full dynamical landscape of hybridization/dehybridization accounting for the sequence-dependent metastable states and association/dissociation pathways remains lacking and fundamental questions remain unresolved. For example, it remains unclear the extent to which hybridization of short DNA oligomers largely proceeds in a conventionally assumed ``all-or-nothing'' fashion or if long-lived metastable states facilitate the transition.\citep{Araque2016LatticeCooperativity, Sikora2013ModelingIntermediates, Wyer2014KineticsAT-tracts, Sanstead2016, Sanstead2018DirectDehybridization, Xiao2019} Out-of-register ``shifted'' base paired structures \citep{Hinckley2014Coarse-grainedEffects, Maciejczyk2014DNAModel, Araque2016LatticeCooperativity, Xiao2019, Romano2013DNADependence,Flamm2000RNAResolution} and frayed structures \citep{Zgarbova2014BaseRNA, Nonin1995TerminalFraying, Nikolova2012ProbingSimulations, Andreatta2006UltrafastHelix} stand as candidates for metastable states with the potential to mediate substantial deviations from all-or-nothing behavior, but the degree to which these states are kinetically relevant is difficult to determine experimentally and is likely to be highly sequence-dependent. The development of predictive models and design rules with which to engineer DNA strands with tailored hybridization/dehybridization kinetics and pathways is vital to advancing rational design of DNA strands for nanotechnology applications and is also of importance in understanding fundamental biological processes such as transcription and gene regulation.

Our understanding of hybridization dynamics is built upon decades of experiments -- such as temperature-jump, salt-jump, pH-jump, and other perturbative methods -- that rapidly stimulate DNA and monitor relaxation to a new equilibrium.\citep{Morrison1993SensitiveSolution, Wetmur1968KineticsDNA, Craig1971RelaxationOligonucleotides, Porschke1973ThermodynamicsPairs, Williams1989LaserDGCATGC, Narayanan2012ExploringMixing, Chen2007InfluenceHybridization, Sanstead2018DirectDehybridization} More recently, single molecule diffusion and tethered multifluorophore assays have facilitated analyses under equilibrium conditions, but these results can be hampered by slow data collection rates and fluorescent tags effects on strand dynamics, particularly for shorter oligomers.\citep{Liu20173DSolution,  Schickinger2018TetheredHelices, Chen2008Base-by-baseSpectroscopy, Dupuis2013Single-moleculeHelices} A number of computational modeling approaches have also been employed to provide molecular-level resolution of hybridization. Simplified lattice models can recapitulate the essential aspects of the hybridization pathways but lack the realism of continuous space representations \citep{Araque2016LatticeCooperativity, Phys2019}. The long time scales associated with hybridization/dehybridization events place them outside the reach of unbiased all-atom molecular dynamics simulations,\citep{Hinckley2014Coarse-grainedEffects} but they can be observed by employing enhanced sampling techniques\citep{Piana2007AtomisticTransition, Zerze2021ThermodynamicsSimulations, Hinckley2013AnHybridization, Schmitt2013ExploringSurface, Sambriski2009, Sambriski2009SequencePathways, Hoefert2011MolecularOligonucleotides,Romano2013DNADependence, Hinckley2014Coarse-grainedEffects} or by using elevated temperature or denaturing solvent concentrations to induce one-way dissociation events.\citep{Wong2008TheSimulations, Perez2010Real-timeUnfolding} The effect of the applied bias upon the thermodynamics can be rigorously corrected for using standard reweighting techniques \citep{Gallicchio2005TemperaturePaths, Souaille2001ExtensionCalculations, Shirts2008StatisticallyStates, Kumar1992THEMethod}. \rood{Rigorous elimination of the bias in the kinetics is critical for the construction of robust and reliable kinetic models that reflect the true system dynamics and are uncontaminated by any residual effects of the biasing potentials used to induce good sampling. A number of approaches to correct the kinetics are also available, including Girsanov reweighting, transition-based reweighting analysis (TRAM), dynamic histogram analysis method (DHAM), and their derivatives~\citep{Prinz2011OptimalDynamics, Chodera2011DynamicalTemperatures, Stelzl2017DynamicSimulations, Donati2017GirsanovModels, Donati2018GirsanovSimulations, Quer2018ANCOORDINATES} \citep{Mey2014XTRAM:States}. The application of these methods under the conditions of high bias necessary for good sampling can, however, present challenges for numerical convergence.} A number of coarse-grained DNA force fields have been developed that enable direct observation of these events over microsecond time scales via unbiased coarse-grained molecular dynamics simulations, \citep{Romano2013DNADependence, Hinckley2013AnHybridization, Maciejczyk2014DNAModel, Markegard2015, Dans2016MultiscaleDNA} which, up to an acceleration factor associated with the smoothing of the underlying free energy landscape inherent to the coarse-graining procedure, can preserve a faithful model of the unbiased dynamics and associated pathways. These models have previously been used to study biological phenomena such as nucleosome dynamics\citep{Lequieu2016Tension-dependentUnwrapping, Lequieu2017InSliding} and transcription factor binding \cite{Terakawa2015P53Simulations, Tan2018DynamicData} as well as nanotechnology applications such as strand displacement \citep{Srinivas2013OnDisplacement, Haley2020DesignDisplacement} and DNA origami \citep{Snodin2019Coarse-grainedOrigami, Doye2020TheOrigami}. \rood{In this study, we choose to employ a coarse-grained model for DNA \citep{Hinckley2013AnHybridization} that is sufficiently inexpensive to enable the collection of sufficient volumes of unbiased simulation trajectories and adequately sample configurational space that we do not need to appeal to biasing strategies to enhance convergence nor apply any \textit{post hoc} corrections to the thermodynamics or kinetics.}

@@

All calculations were performed using the LAMMPS simulation package (\url{http://lammps.sandia.gov}) in accordance with best practices for the 3SPN.2 model. \citep{Plimpton1997FastDynamics} A single pair of self-complementary sequences were placed in a cubic periodic box with side length 7.8 nm corresponding to a single-strand concentration of \rood{7 mM}. \rood{This concentration is only 3.5$\times$ larger than the 2 mM concentration employed in our experimental analyses (Section 2.2).} Solvent effects were modeled implicitly by employing Langevin dynamics \citep{DunwegB.Paul1991BrownianNumbers., Bussi2007AccurateDynamics} with an experimentally motivated per-site friction coefficient of 9.94 $\times$ 10$^{-11}$ m$^2$/s. \citep{Nkodo2001DiffusionElectrophoresis, Hinckley2013AnHybridization} We specified a 240 mM implicit salt concentration and treated electrostatic interactions using the Debye-H{\"u}ckel with a 5 nm cutoff radius. \citep{Debye1923ZurElektrolyte} Simulations were performed in the NVT ensemble employing a Langevin thermostat. \citep{Schneider1978Molecular-dynamicsTransitions}. \rood{With the exception of the simulation data reported in Section 3.1 where we draw comparisons against experimental data collected over a range of temperatures,} each sequence was simulated at its \rood{respective melting temperature dictated by the 3SPN.2 model} -- AT-all: 309 K, GC-end: 317 K, GC-core: 324 K, GC-mix: 324 K -- in order to maximize the number of spontaneous transitions between dissociated and hybridized states. Melting temperatures for each sequence were determined empirically by comparing the ratio of hybridized to dissociated populations over a 10 K temperature ramp centered on the nearest neighbor (NN) model predicted melting temperature,~\citep{SantaLucia1998AThermodynamics, Santalucia2004TheMotifs} and selecting the temperature at which the hybridized and dissociated populations were approximately equal. The Langevin equations of motion were integrated using the scheme of Bussi and Parrinello \citep{Bussi2007AccurateDynamics} with a 20 fs integration time step. We performed 40 independent simulations for each of the four sequences with half of the runs initialized from the hybridized state and half from the dissociated state. The initial hybridized state was defined based on the crystal structure coordinates of Arnott et al.~\citep{Arnott1976CRCBiology} The dissociated state was generated from the hybridized state by displacing one strand away from the other by 1 nm in each of the $x$, $y$, and $z$ directions. The magnitude of this displacement did not affect the results so long as all native Watson-Crick (WC) bonds were completely broken. Initial bead velocities were assigned from a Maxwell-Boltzmann distribution at the temperature of interest. Each simulation was conducted for 26 $\mu$s and frames saved to disc every 100 ps. Each simulation required $\sim$24 CPU-hours on 28$\times$Intel E5-2680v4 CPU cores. The first 1 $\mu$s of each run was discarded for equilibration providing us with 40$\times$25 $\mu$s = 1 ms of simulation data for each sequence, during which time we observed 55-100 hybridization/dehybridization events.

@@

Markov state models (MSMs) are a powerful approach to infer long-time kinetic models from short molecular simulation trajectories \citep{Sengupta2019AutomatedSelf-assembly, Pande2010EverythingAsk, Chodera2014MarkovDynamics, Husic2018MarkovScience, Wehmeyer2019IntroductionSoftware} that we employ in this work to construct high-resolution sequence-dependent kinetic models of DNA hybridization and dissociation. In brief, a MSM extracts from simulation trajectories an ensemble of long-lived metastable macrostates, their equilibrium occupancy populations, and the equilibrium probability fluxes between them. In this manner, they provide an interpretable and predictive model of the system thermodynamics and kinetics. MSMs have recently been implemented to study the hybridization mechanism of one particular 14-mer DNA oligomer, but determining the sequence-dependent kinetics and mechanisms was not the focus of this study.\citep{Jin2019} An energy disconnectivity graph-based approach was used to interrogate the differences in hybridization rates and mechanisms between GGGGGG and GCGCGC hexamers to reveal strong deviations from ``all-or-nothing'' behaviors and the importance of zippering and \rood{out-of-register diffusion mechanisms}.~\citep{Xiao2019} Kinetic models were constructed not from the dynamical trajectories of the molecular model, but by estimating rate constants between local minima using a transition state theory approximation. A recent application of MSMs to the long-time dynamics of short RNA oligonucleotides revealed stacking time scales to be highly sequence dependent.~\citep{Pinamonti2017} In this work, MSMs were constructed for each of the four DNA sequences \rood{at their respective 3SPN.2 melting temperatures} from the 40$\times$25 $\mu$s simulation trajectories following a four-step protocol detailed in Ref.~\citep{Sidky2019High-ResolutionVAMPnets}: (i) trajectory featurization, (ii) dimensionality reduction, (iii) microstate clustering and microstate transition matrix inference, (iv) macrostate clustering and macrostate transition matrix inference. Calculations were performed using the PyEMMA software package. \citep{Scherer2015PyEMMAModels}

@@

We first sought to demonstrate that the coarse-grained 3SPN.2 model accurately recapitulates the experimentally observed kinetics of DNA oligomer hybridization/dissociation by validating our computational predictions against temperature-jump infrared (T-jump IR) experiments. We conducted T-jump IR experiments for each DNA sequence as a function of temperature and extracted \rood{the ``slow'' $k_d^\mathrm{slow}$  and ``fast'' $k_d^\mathrm{fast}$ rates, corresponding to processes proceeding on 10-30 $\mu$s and 70-100 ns timescales, respectively}. The slow response has previously been attributed to duplex dissociation on microsecond time scales induced by the rapid heating of the initially hybridized duplex.~\citep{Sanstead2016, Sanstead2018DirectDehybridization} The fast response has been assigned to terminal base pair fraying,~\citep{Sanstead2016, Sanstead2018DirectDehybridization} with this process corresponding to a relatively complex dynamical process that can span time scales from picoseconds to microseconds.~\cite{Nonin1995TerminalFraying, Nikolova2012ProbingSimulations, Andreatta2006UltrafastHelix, Galindo-Murillo2015ConvergenceDGCACGAACGAACGAACGC} All-atom simulations suggest that frayed ends can assume misaligned WC bonds, base-sugar hydrogen bonds, and terminal stacked conformations.~\citep{Pinamonti2019TheModels, Zgarbova2014BaseRNA}

To computationally mimic the T-jump process in our 3SPN.2 simulations, we conducted 1 $\mu$s simulations of an initially hybridized DNA duplex over a range of temperatures and monitored its structural relaxation. We performed 120 independent simulations for each DNA sequence at each temperature, and from these extracted computational estimates of $k_d^\mathrm{slow}$ and $k_d^\mathrm{fast}$ (\blauw{Fig.~S4}). First, we tracked the slow response corresponding to duplex dissociation in our simulations by compiling the distribution of times at which both of the central base pairs first separate to a distance of 1.3 nm starting from an initial fully hybridized duplex. This cutoff was selected as the distance beyond \rood{which} the strands are effectively non-interacting and defines the dissociated state. \rood{We verified that our results were robust to choices of this cutoff over the range 0.8-2.5 nm.} We extracted our computational estimate of $k_d^\mathrm{slow}$ by fitting a decaying exponential to the fraction of hybridized sequences as a function of time $f_\mathrm{hybridized}(t) = \exp(-k_d^\mathrm{slow}t)$. \rood{To correct for finite-size effects associated with the simulation of a single duplex within a periodic simulation box and ensure that we will observe asymptotically correct behavior at long times, we apply the fluctuation correction proposed by Ouldridge et al.\citep{Ouldridge2010ExtractingSimulations} to our predictions for $f_\mathrm{hybridized}$ bulk yields. We correct the values observed in our simulations $f_\mathrm{hybridized}^\mathrm{sim}$ to the bulk corrected result $f_\mathrm{hybridized}$ as,}

    \begin{equation}\label{}
    \phi = \frac{f_\mathrm{hybridized}^\mathrm{sim}}{1 - f_\mathrm{hybridized}^\mathrm{sim}}
    \end{equation}

    \begin{equation}\label{}
	f_\mathrm{hybridized} = (1+\frac{1}{4\phi}) - \sqrt{(1+\frac{1}{4\phi})^2 - 1}
	\end{equation}

Second, we tracked the fast response corresponding to terminal base pair fraying by compiling the distribution of times at which either of the terminal base pairs first separated to a distance of 1.3 nm, corresponding to a complete breakage of the WC interaction. We extracted our computational estimate of $k_d^\mathrm{fast}$ through a decaying exponential fit to the fraction of unfrayed sequences as a function of time $f_{\mathrm{unfrayed}}(t) = \exp(-k_d^\mathrm{fast}t)$.

We present in Fig.~\ref{fig:relaxation-comparison} a comparison of $k_d^\mathrm{slow}$ and $k_d^\mathrm{fast}$ estimated by computation and experiment. Although the 3SPN.2 model was not directly fitted against kinetic data,~\cite{Hinckley2013AnHybridization} its predictions of sequence dependent T-jump relaxation rates are, within a systematic scaling factor in time and systematic shift in temperature, in good agreement with observed experimental trends. It is well-known that the smoothing of the underlying free energy landscape induced by coarse-graining artificially accelerates the kinetics of coarse-grained molecular simulations and that different degrees of freedom may be accelerated by different factors.~\citep{Marrink2007TheSimulations, Fritz2011MultiscaleDynamics, Marrink2013PerspectiveModel} We find that the simulated slow responses corresponding to center-of-mass translation of the strands during dissociation of the duplex is $\sim$10$\times$ accelerated relative to experiment, whereas the fast responses corresponding to fraying of the terminal bases is $\sim$120$\times$ accelerated. We apply these sequence-independent scaling factors to our reported computational values in Fig.~\ref{fig:relaxation-comparison}. Although the 3SPN.2 model reproduces melting temperatures relatively well, we observed a systematic 4 K under-prediction relative to experiment and so we apply a \rood{universal (+4) K corrective temperature shift} to our computational results. We note that these empirical calibration factors to the 3SPN.2 predictions are applied only for the purposes of making an experimental comparison, but acknowledge that there are uncertainties introduced by assuming that the equilibrium dynamics at fixed temperature can be compared directly to relaxation kinetics following a 15$^\circ$C T-jump. The computational time scales and melting temperatures reported in the remainder of the paper are not corrected \rood{by these calibration corrections or finite-size fluctuation corrections} since we are only interested in the relative trends in the behaviors of the four sequences.
